Article Title

### 2014-05-05 plos.tex
completed figures 3, 4, 5, and 6.

Given an arbitrary covariance matrix estimate $C$, the corresponding correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}\label{eq:precision}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2},
\end{equation}
where $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$:
\begin{equation}\label{eq:partial}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}

To see that this is the case, partition the covariance matrix 
\begin{equation}
C = \begin{pmatrix}
C_{11} & C_{12} \\
C_{12}^\T & C_{22}
\end{pmatrix}
\end{equation}
so that $C_{11}$ is the $2\times 2$ covariance matrix of two cells of interest. Such partitioning can be done for any pair of cells by changing their order. 
Then the covariance matrix of the conditional distribution of these two cells is \cite{Anderson:2003}:
\begin{equation}
\hat C_{11} = C_{11}-C_{12}C^{-1}C_{12}^\T 
\end{equation}
The inverse of the conditional covariance matrix can be found in the corresponding partition of the precision matrix $C^{-1}$:
\begin{equation}
C^{-1}=\begin{pmatrix}
\hat C_{11}^{-1} &   -\hat C_{11}^{-1}C_{12}C_{22}^{-1} \\
-C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}  &  C_{22}^{-1}+C_{22}^{-1}C_{12}^\T\hat C_{11}^{-1}C_{12}C_{22}^{-1}
\end{pmatrix}
\end{equation}
The inverse of a $2\times 2$ correlation matrix reverses the sign of its off-diagonal elements:
\begin{equation}
\begin{pmatrix}
1 & \rho \\
\rho & 1
\end{pmatrix}^{-1}
\propto 
\begin{pmatrix}
-1 & \rho \\
\rho & -1 
\end{pmatrix}
\end{equation}
Therefore, scaling $C^{-1}$ to have 1s on the diagonal yields the partial correlation in each of its $2\times 2$ partitions, resulting in a matrix of partial pairwise correlations.

@@

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
    {\bf Row 1.} Graphical representations of the target estimates of the four respective regularized covariance matrix estimators.  Recorded neurons are represented by green spheres and latent units by light-shaded spheres.  Edges represent non-zero partial correlations, \ie \sq{interactions}.
        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies. 
        {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units. 
        {\bf  Row 1, C}. For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units. 
        {\bf  Row 1, D}.  For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} with several latent units. \\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices shown in Row 2.  The structure of the sample correlation matrix is difficult to discern.
\\
{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3. The regularized estimates are closer to the truth than the sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) for the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with structure that matches the true model converged to zero faster than the other estimators. Some exceptions occur when the sample is too small to reveal the structure in the data. In these cases estimators with simpler structures can be more efficient than estimators with true structure.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) for the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:1}
\end{FPfigure}

@@

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}). 
    {\bf C--G.} Calcium activity data from an example site.
    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.  The traces from {\bf C} are highlighted in red.
    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site. Only spiking cells are shown. The cells' colors indicate their orientation preferences. The gray cells were not significantly tuned.
    {\bf F.} The sample noise correlation matrix of the activity of the neural population. 
    {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
} \label{fig:2}}
    {\includegraphics[width=8.3cm]{figures/Figure03.pdf}}
\end{figure}

@@

To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Methods for details). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}). The respective median differences of the cross-validation loss were 0.018, 0.0025, 0.0029, and 0.0017 nats/cell/bin, significantly greater than zero ($p<10^{-5}$ in each comparison, Wilcoxon signed rank test).

@@

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.} 
    Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:2} and \ref{fig:4} is circled in blue. 
\\
    {\bf A.} The number of inferred latent units vs.~population size.
    {\bf B.} The average node degree for sparse partial correlations vs.~population size.
    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
    {\bf E.} Percent overlap between the graph of highest sample correlations and the graph of interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component. 
    {\bf F.} The percentage of negative interactions vs.~connectivity in the $C_{\sf sparse+latent}$ estimates.
}
\label{fig:5}}
{\includegraphics{./figures/Figure06.pdf}}
\end{figure}

@@

\begin{figure}[!ht]
    \begin{center}
        \includegraphics{./figures/Figure07.pdf}
    \end{center}
    \caption{{\bf Dependence of sample correlations and regularized partial correlations on orientation tuning and physical distance.}
    {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites. At each site the correlations were normalized by the respective average correlation shown in \figref{5}{D}.  The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel. Sites were included if they had at least 20 pairs of neurons in each of the intervals. The error bars indicate the standard error of the mean based on on the number of sites $n$.
    {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$25$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0--25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
    {\bf D--F.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component obtained from $C_{\sf sparse+latent}$. Normalized connectivity was computed as the fraction of pairs connected by interactions of corresponding signs in each interval divided by the fraction of non-zero interactions across the entire site. }
\label{fig:6}
\end{figure}

@@

\begin{FPfigure}
\begin{center}
\includegraphics{./figures/Figure-Supp01.pdf}
\end{center}
\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators with respect to multivariate normal cross-validation loss.}
\\{\bf A--D.} These panels are identical to those in \figref{3}{A--D}.
\\{\bf E.} $C_{\sf sparse}$ outperforms $C_{\sf sample}$: median improvement 0.15 nats/neuron, $p=3.9\times 10^{-5}$.
\\{\bf F.} $C_{\sf sparse}$ performs similarly to $C_{\sf diag}$: median improvement $1.0\times10^{-3}$ nats/neuron, $p=0.13$.
\\{\bf G.} $C_{\sf sparse}$ outperforms $C_{\sf factor}$: median improvement $0.15$ nats/neuron, $p=1.2\times10^{-4}$.
\\{\bf H.} $C_{\sf factor}$ outperforms $C_{\sf sample}$: median improvement $0.029$ nats/neuron, $p=2.7\times10^{-4}$.
\\{\bf I.} $C_{\sf factor}$ performs worse than $C_{\sf diag}$: median improvement $-0.16$ nats/neuron, $p=1.6\times10^{-4}$.
\\{\bf J.} $C_{\sf diag}$ outperforms $C_{\sf sample}$: median improvement $0.14$ nats/neuron, $p=2.7\times10^{-5}$.
}
\label{supp:01}
\end{FPfigure}

\begin{figure}[!ht]
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
{\caption{{\bf The sparse+latent estimator outperforms the other estimators under the quadratic loss (Eq.~\ref{eq:quadratic}) just as it did under the normal loss (Eq.~\ref{eq:loss}) used in Fig.~\ref{fig:3}.}
    Histograms of the cross-validation loss of estimators $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ using the quadratic loss function.
    The histograms were obtained from 31 imaged sites in 15 mice. 
    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
    The arrow heads indicate the results for the site shown in Fig.~\ref{fig:2} and Fig.~\ref{fig:4}.
    The $p$-values are for the two-sided Wilcoxon's signed rank test.
    \\{\bf A.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sample}$: median $\Delta\ell_{C_{\sf sample},C_{\sf sparse+latent}}$ = 0.35, $p=1.7\times 10^{-6}$.
    \\{\bf B.} $C_{\sf sparse+latent}$ outperforms $C_{\sf diag}$: median $\Delta\ell_{C_{\sf diag},C_{\sf sparse+latent}}$ = $2.5\times 10^{-3}$, $p=1.0\times 10^{-4}$.
    \\{\bf C.} $C_{\sf sparse+latent}$ outperforms $C_{\sf factor}$: median $\Delta\ell_{C_{\sf factor},C_{\sf sparse+latent}}$ = 0.27, $p=1.2\times 10^{-6}$.
    \\{\bf D.} $C_{\sf sparse+latent}$ outperforms $C_{\sf sparse}$: median $\Delta\ell_{C_{\sf sparse},C_{\sf sparse+latent}}$ = $3.7\times 10^{-4}$, $p=2.2\times 10^{-3}$.
}
\label{supp:02}}
{\includegraphics{./figures/Figure-Supp02.pdf}}
\end{figure}

>>>>

Given an arbitrary covariance matrix estimate $C$, the corresponding correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}\label{eq:precision}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2},
\end{equation}
where $\circ$ denotes the entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. 
Similarly, the matrix of partial correlations $P$ is computed by normalizing the \emph{precision matrix} $C^{-1}$ \cite{Whittaker:2009}:
\begin{equation}\label{eq:partial}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}

@@

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure01.pdf}
    \end{center}
    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
    {\bf Row 1.} Graphical representations of the target estimates of the four respective regularized covariance matrix estimators.  Recorded neurons are represented by green spheres and latent units by light-shaded spheres.  Edges represent non-zero partial correlations, \ie \sq{interactions}.
        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies. 
        {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units. 
        {\bf  Row 1, C}. For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units. 
        {\bf  Row 1, D}.  For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} with several latent units. \\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices shown in Row 2.  The structure of the sample correlation matrix is difficult to discern.
\\
{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3. The regularized estimates are closer to the truth than the sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) for the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with structure that matches the true model converged to zero faster than the other estimators. Some exceptions occur when the sample is too small to reveal the structure in the data. In these cases estimators with simpler structures can be more efficient than estimators with true structure.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) for the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:1}
\end{FPfigure}

@@

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Acquisition of neural signals for the estimation of noise correlations.}
    Visual stimuli comprising full-field drifting gratings interleaved with blank screens ({\bf A}) presented during two-photon recordings of somatic calcium signals using fast 3D random-access microscopy ({\bf B}). 
    {\bf C--G.} Calcium activity data from an example site.
    {\bf C.} Representative calcium signals from eight cells out of 298 cells downsampled to 20 Hz. The inferred firing rate binned in 150 ms intervals are indicated by red ticks below each trace.
    {\bf D.} The raster plot of the inferred firing rates, binned in 150 ms intervals, from the entire population from the first (left) and last (right) minute of the entire recording.  The traces from {\bf C} are highlighted in red.
    {\bf E.} The spatial arrangement and orientation tuning of the 298 cells from the imaged site. Only spiking cells are shown. The cells' colors indicate their orientation preferences. The gray cells were not significantly tuned.
    {\bf F.} The sample noise correlation matrix of the activity of the neural population. 
    {\bf G.} Histogram of noise correlation coefficients. The red line indicates the mean.
} \label{fig:2}}
    {\includegraphics[width=8.3cm]{figures/Figure02.pdf}}
\end{figure}

@@

To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Methods for details). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}). The respective median differences of the cross-validation loss were 0.019, 0.0027, 0.0031, and 0.0022 nats/cell/bin, significantly greater than zero ($p<10^{-4}$ in each comparison, Wilcoxon signed rank test).

@@

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.} 
    Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:2} and \ref{fig:4} is circled in blue. 
\\
    {\bf A.} The number of inferred latent units vs.~population size.
    {\bf B.} The average node degree for sparse partial correlations vs.~population size.
    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
    {\bf E.} Percent overlap between the graph of highest sample correlations and the graph of interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component. 
    {\bf F.} The percentage of negative interactions vs.~connectivity in the $C_{\sf sparse+latent}$ estimates.
}
\label{fig:5}}
{\includegraphics{./figures/Figure05.pdf}}
\end{figure}

@@

\begin{figure}[!ht]
    \begin{center}
        \includegraphics{./figures/Figure06.pdf}
    \end{center}
    \caption{{\bf Dependence of sample correlations and regularized partial correlations on orientation tuning and physical distance.}
    {\bf A--C.} Average partial correlations (red) estimated by $C_{\sf sparse+latent}$ and average sample correlations (black) averaged across multiple imaged sites. At each site the correlations were normalized by the respective average correlation shown in \figref{5}{D}.  The number of sites, $n$, that qualified to be included in the analysis is indicated in each panel. Sites were included if they had at least 20 pairs of neurons in each of the intervals. The error bars indicate the standard error of the mean based on on the number of sites $n$.
    {\bf A.} Average correlations between pairs of neurons tuned to orientation with differences in preferred orientation in the intervals of 0--15$^\circ$, 15--45$^\circ$ and 45--90$^\circ$. 
    {\bf B.} Average correlations between pairs of neurons located at the same depth ($\pm$25$\mu$m) separated by lateral distances in the intervals of 0--25 $\mu$m, 25--75 $\mu$m, 75--150 $\mu$m, and 150+ $\mu$m.
    {\bf C.} Average correlations between pairs of neurons displaced laterally by less than 25 $\mu$m separated in depth by distances in the intervals of 0--25 $\mu$m, 25--60 $\mu$m, and 60+ $\mu$m.
    {\bf D--F.} Normalized connectivity of positive (green) and negative (dark red) interactions from the sparse component obtained from $C_{\sf sparse+latent}$. Normalized connectivity was computed as the fraction of pairs connected by interactions of corresponding signs in each interval divided by the fraction of non-zero interactions across the entire site. }
\label{fig:6}
\end{figure}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/src/Supp1.eps}
\end{center}
\caption{Comparison of the $C_{\sf sparse}$ estimator to the other estimators with respect to the normal loss function.}
\label{supp:01}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/src/Supp2.eps}
\end{center}
\caption{Comparison of the $C_{\sf sparse+latent}$ estimator to the other estimators with respect to the quadratic loss.}
\label{supp:02}
\end{figure}
